
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Dedicatory

\chapter*{}
\pagenumbering{Roman} 
\begin{flushright}
\textit{Dedicated to \\
my family, partner and friends.}
\end{flushright}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Acknowledgements

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements} 
\markboth{Acknowledgements}{Acknowledgements} 

With this work, I finalize one of the most beautiful and difficult periods of my life, my stage as a student. I have achieved this great challenge, first of all, thanks to my family, my parents and my brother. They have supported me since I was a child, helping me in all my decisions and doing all their efforts unconditionally. Thanks to them today I am an engineer.  

Thanks to my partner, always by my side. He has accompanied me during all these years, being my best support, colleague, friend and love. I could not have chosen better, I love you.

Thanks to my friends, my girls since we were three years old. We have grown together, living and enjoying each stage of our lives. A friend is a treasure, so I am the luckiest person in the world with you. 

I also want to thank my schoolmates during the degree. We started by helping each other in the battle and, at the end, you have become my friends. Thanks for helping me in the worst moments.

I would like to thank all the Dr. Scratch team. Especially, thanks to Cristian, who helped me during these two years and taught me so many things.

Finally, I could not have developed this project without my tutor, Gregorio. Thanks for this great opportunity and for all the things that I have learned from it.

To all of you: Thank you.




